% !TEX root = ../../main.tex
\section{Streaming Pipeline: Eventsim → Kafka → Flink → Snowflake}

This section walks through the four active stages of the real-time pipeline:
event generation, message transport, stream processing, and observability.

\subsection{Event Simulation with Eventsim}

Eventsim is an open-source event generator that simulates the behaviour of users on a
fictional music streaming website.
It models realistic user sessions — login, browse, play tracks, skip songs, logout —
and emits each interaction as a JSON event to a Kafka topic.

\begin{itemize}
  \item Song metadata is drawn from a \textbf{10,000-song subset} of the Million Songs Dataset,
        providing realistic artist names, song titles, and durations.
  \item User demographics (age, location, subscription tier) are synthetically generated
        using US census data, giving the events geographic and behavioural diversity.
  \item Three topics are produced simultaneously:
    \begin{itemize}
      \item \texttt{listen\_events} — artist, song, duration, timestamp, user ID, location, subscription level.
      \item \texttt{page\_view\_events} — page name, HTTP method, status code, session ID, user context.
      \item \texttt{auth\_events} — session ID, success/failure flag, subscription level, timestamp.
    \end{itemize}
  \item Eventsim~\cite{eventsim} runs as a Docker Compose~\cite{docker_compose} service that starts automatically after the Kafka topics
        are created by \texttt{kafka-init}, using \texttt{depends\_on: service\_completed\_successfully}
        to enforce ordering.
  \item The simulation window is set to \texttt{start = now}, \texttt{end = now + 24 hours} using
        GNU \texttt{date} inside the container, so the event timestamps are always anchored to
        the current wall-clock time.
\end{itemize}

\subsection{Apache Kafka: 3-Node KRaft Cluster}

The project uses \textbf{Apache Kafka 4.3.0}~\cite{kafka} running in \textbf{KRaft mode}~\cite{kafka_kraft} — the native Kafka
consensus protocol that eliminates the Zookeeper dependency introduced in earlier versions.
The cluster consists of three nodes (\texttt{broker-1}, \texttt{broker-2}, \texttt{broker-3}),
each acting simultaneously as both broker and KRaft controller.

Key cluster configuration:
\begin{itemize}
  \item \textbf{Replication factor 3, min ISR 2} — every partition has a replica on each broker,
        so the cluster tolerates one broker failure without data loss or write interruption.
  \item \textbf{3 partitions per topic} — each of the three event topics
        (\texttt{listen\_events}, \texttt{page\_view\_events}, \texttt{auth\_events}) is
        pre-created with three partitions by the \texttt{kafka-init} service at startup.
        This matches the Flink parallelism of 3, allowing each TaskManager to own exactly one
        partition of each topic.
  \item \textbf{Fixed \texttt{CLUSTER\_ID}} — a static cluster ID ensures all three broker
        containers format the same KRaft metadata log on every \texttt{compose up}, preventing
        cluster-ID mismatch errors across restarts.
\end{itemize}

\subsection{Flink Streaming Job: Scaled Deployment}

The Flink cluster is scaled to \textbf{three TaskManagers}, each allocated 1,728\,MB of process
memory and four task slots.
The PyFlink job runs at \textbf{parallelism 3}, and with Flink's default slot-sharing enabled,
each TaskManager executes one complete Source $\to$ UDF $\to$ Sink pipeline — consuming one
partition of each topic in parallel.
The \texttt{cluster.evenly-spread-out-slots: true} option ensures the three task chains are
distributed evenly across all three TaskManagers rather than being packed onto one.

\textbf{Checkpointing}~\cite{flink_checkpointing} is enabled with a 10-second interval
(\texttt{env.enable\_checkpointing(10000)}).
This is essential for two reasons: it provides fault tolerance by snapshotting Flink operator
state to durable storage, and it causes the Kafka source to commit consumer-group offsets back
to Kafka on each checkpoint — making the \texttt{flink-streamify} consumer group visible to
\texttt{kafka-exporter} and therefore to the monitoring dashboards.

\textbf{Auto-submission} is handled by a dedicated \texttt{flink-job-submitter} service that
waits until at least three free task slots are available, checks whether a job is already
\texttt{RUNNING} to avoid duplicates on repeated \texttt{compose up} calls, and then submits
the PyFlink script in detached mode (\texttt{flink run -d}).

The PyFlink job~\cite{flink_pyflink} (\texttt{flink\_streaming/jobs/stream\_events.py}) consumes all three topics in
a single Flink execution graph using the Table API and Flink SQL~\cite{flink_kafka_connector}.

\subsubsection{Schema and Source Registration}

Each topic's JSON payload is mapped to a set of Flink SQL column definitions in \texttt{schema.py}.
The \texttt{create\_kafka\_source()} function registers a temporary Kafka source table via DDL:

\begin{lstlisting}[language=SQL, caption={Generated Kafka source DDL for \texttt{listen\_events}.}]
CREATE TEMPORARY TABLE src_listen_events (
    `artist`        STRING,
    `song`          STRING,
    `duration`      DOUBLE,
    `ts`            BIGINT,
    `userId`        BIGINT,
    `level`         STRING,
    `city`          STRING,
    `state`         STRING,
    -- ...
) WITH (
    'connector'                    = 'kafka',
    'topic'                        = 'listen_events',
    'properties.bootstrap.servers' = 'broker:29092',
    'properties.group.id'          = 'flink-streamify',
    'scan.startup.mode'            = 'earliest-offset',
    'format'                       = 'json',
    'json.ignore-parse-errors'     = 'true'
)
\end{lstlisting}

\subsubsection{Transformation and Sink}

For each topic, the job enriches the raw stream and writes to a Snowflake JDBC sink:

\begin{itemize}
  \item \textbf{Timestamp conversion} — the raw \texttt{ts} field (epoch milliseconds) is converted
        to a proper \texttt{TIMESTAMP(3)} via \texttt{TO\_TIMESTAMP\_LTZ(ts / 1000, 3)}.
  \item \textbf{Partition columns} — \texttt{year\_}, \texttt{month\_}, \texttt{day\_}, \texttt{hour\_}
        are derived from the timestamp for efficient Snowflake partitioning.
  \item \textbf{String decoding UDF} — artist and song names in \texttt{listen\_events} and
        \texttt{page\_view\_events} are stored in a latin1 unicode-escape encoding.
        A Python UDF \texttt{string\_decode} converts them to UTF-8 before writing.
  \item \textbf{Processing timestamp} — \texttt{CURRENT\_TIMESTAMP} is appended as \texttt{processed\_time}
        to track pipeline latency.
\end{itemize}

All three INSERT statements are registered in a single \texttt{StatementSet} and submitted as
\textbf{one Flink job}, sharing one checkpoint cycle and one set of TaskManager slots:

\begin{lstlisting}[language=bash, caption={Submitting the PyFlink job after \texttt{docker compose up}.}]
docker exec -it flink-jobmanager \
    flink run -py /opt/flink/jobs/stream_events.py
\end{lstlisting}
