\section{Kafka in the Streamify Pipeline}

Streamify deploys a \textbf{Confluent Platform} Kafka broker inside Docker Compose,
leveraging the Confluent image (\texttt{confluentinc/cp-kafka}) alongside the Confluent
Schema Registry and Confluent Control Center for schema enforcement and real-time monitoring.

\subsection{Broker Configuration}

The broker is configured with two listener addresses:

\begin{itemize}
  \item \textbf{\texttt{PLAINTEXT://broker:29092}} — the internal Docker network listener
        used by all other containers (Eventsim producer, PyFlink consumer, Control Center).
  \item \textbf{\texttt{PLAINTEXT\_HOST://localhost:9092}} — the host machine listener,
        allowing local development tools (e.g.\ \texttt{kafka-console-consumer}) to connect.
\end{itemize}

Key broker settings relevant to the pipeline:

\begin{table}[H]
  \centering
  \caption{Confluent Kafka broker settings in Streamify.}
  \label{tab:kafka-config}
  \begin{tabularx}{\textwidth}{lX}
    \toprule
    \textbf{Setting}                         & \textbf{Value / Effect} \\
    \midrule
    \texttt{KAFKA\_OFFSETS\_TOPIC\_REPLICATION\_FACTOR} & 1 — single-broker local deployment; no replication needed. \\
    \texttt{KAFKA\_GROUP\_INITIAL\_REBALANCE\_DELAY\_MS} & 0 — consumer groups rebalance immediately on startup, avoiding cold-start delays. \\
    \texttt{KAFKA\_TRANSACTION\_STATE\_LOG\_MIN\_ISR}    & 1 — allows exactly-once transactions with a single broker. \\
    \texttt{KAFKA\_TRANSACTION\_STATE\_LOG\_REPLICATION\_FACTOR} & 1 — consistent with single-broker setup. \\
    \bottomrule
  \end{tabularx}
\end{table}

\subsection{Topics}

Three topics are auto-created by Eventsim on first publish.
Each topic is configured with \textbf{1 partition} (sufficient for the single-consumer Flink job)
and \textbf{1 replica} (single-broker deployment):

\begin{itemize}
  \item \textbf{\texttt{listen\_events}} — one record per song-play interaction.
        Fields include: \texttt{artist}, \texttt{song}, \texttt{duration}, \texttt{ts} (epoch ms),
        \texttt{userId}, \texttt{level} (free/paid), \texttt{city}, \texttt{state}, \texttt{lat}, \texttt{lon}.
  \item \textbf{\texttt{page\_view\_events}} — one record per page navigation.
        Fields include: \texttt{page}, \texttt{method}, \texttt{status}, \texttt{sessionId},
        \texttt{userId}, \texttt{ts}, \texttt{userAgent}.
  \item \textbf{\texttt{auth\_events}} — one record per login/logout.
        Fields include: \texttt{success} (boolean), \texttt{level}, \texttt{sessionId}, \texttt{ts}.
\end{itemize}

All events are serialised as \textbf{JSON} (not Avro) to keep schema management simple
for this educational project.
A Confluent Schema Registry (port \texttt{8081}) is deployed alongside the broker
and is available for future migration to Avro with schema evolution support.

\subsection{Producer: Eventsim}

Eventsim runs in its own Docker container and publishes events to the broker at
\texttt{broker:29092} using the standard Confluent Kafka Java producer.
It simulates a configurable number of concurrent users (\texttt{--nusers}) over a specified
time window (\texttt{--from}, \texttt{--to}), producing a realistic mix of listen, page-view,
and authentication events based on the Million Songs Dataset.

The producer does \emph{not} use a message key, so events are distributed across partitions
round-robin.
Because each topic has only one partition in the local setup, all records land in partition 0
and are therefore \textbf{globally ordered within each topic}.

\subsection{Consumer: Apache Flink (PyFlink)}

The PyFlink job registers all three topics as Flink SQL source tables using the
\texttt{flink-connector-kafka} library.
The consumer is configured with:

\begin{itemize}
  \item \textbf{Consumer group ID}: \texttt{flink-streamify} — a single group shared across
        all three source operators so that Kafka tracks one set of offsets per topic.
  \item \textbf{Startup mode}: \texttt{earliest-offset} — on cold start, Flink reads from the
        beginning of each partition, ensuring no events are missed if Eventsim started before Flink.
  \item \textbf{Exactly-once integration}: Flink's checkpoint barriers coordinate with the
        Kafka consumer's offset commits.
        Offsets are only committed to Kafka once the corresponding Flink checkpoint is
        successfully persisted — preventing any event from being skipped or re-processed
        on recovery.
\end{itemize}

All three source operators are combined in a single \texttt{StatementSet}, which Flink compiles
into one execution graph with a shared checkpoint cycle.
This means all three Kafka→Snowflake pipelines advance their checkpoints together,
guaranteeing a consistent cut across all topics.

\subsection{Monitoring with Confluent Control Center}

The Confluent Control Center (port \texttt{9021}) provides a web UI for observing the live
state of the Kafka cluster:

\begin{itemize}
  \item \textbf{Topic throughput} — messages per second produced and consumed, per topic.
  \item \textbf{Consumer lag} — the offset delta between the latest produced record and the
        last committed consumer offset; a rising lag indicates that Flink is falling behind.
  \item \textbf{Partition leaders} — which broker holds the leader replica for each partition.
  \item \textbf{Schema Registry} — registered schemas and their compatibility levels.
\end{itemize}

In the local Docker Compose setup, the Control Center connects to the broker at
\texttt{broker:29092} and to the Schema Registry at \texttt{schema-registry:8081}.
