\section{Why Apache Flink Was Chosen for Streamify}

After evaluating Apache Spark Structured Streaming and Apache Flink against the requirements
of the Streamify pipeline, Flink was selected as the stream-processing engine.
This section explains the five decisive factors behind that choice.

\subsection{Lower Latency Without Configuration Trade-offs}

Spark's default micro-batch mode imposes an irreducible latency floor equal to the trigger
interval.
Even setting the interval to its minimum (1\,ms) does not eliminate the per-batch planning
overhead — catalyst optimisation, task scheduling, and shuffle materialisation all add latency
before a single record reaches the sink.

Flink's true event-at-a-time model has no such floor.
Each record flows through the operator DAG and reaches the Snowflake JDBC sink within the
time it takes to execute the processing logic for that single record — typically single-digit
milliseconds on a local Docker deployment.
For a music streaming analytics platform where dashboard freshness directly impacts user
experience, this difference is meaningful even in a prototype: shorter end-to-end latency
means analysts see the current play counts rather than those from the last trigger window.

\subsection{Correct Event-time Handling via Watermarks}

Eventsim timestamps events at the moment they are \emph{generated} (event time), not the moment
they arrive at Kafka or Flink (processing time).
Network jitter, Docker scheduling delays, and Eventsim's own batching mean that events can
arrive at Flink slightly out of order relative to their timestamps.

Flink provides \textbf{first-class event-time support} through its watermark mechanism.
A watermark strategy (e.g.\ \texttt{WatermarkStrategy.forBoundedOutOfOrderness(Duration.ofSeconds(5))})
can be attached directly to the Kafka source, telling Flink's windowed operators to wait up to
5 seconds for late-arriving events before closing a window.
In Spark, event-time watermarks are also supported but require the user to explicitly configure
\texttt{withWatermark()} on each streaming DataFrame and are less tightly integrated with the
checkpoint mechanism.
Flink's watermark infrastructure required zero additional code in Streamify: the default
monotonous watermark on the \texttt{ts} column was sufficient for the project's aggregation queries.

\subsection{Single Job for Multiple Sinks via StatementSet}

Streamify must write three independent streams — \texttt{listen\_events},
\texttt{page\_view\_events}, and \texttt{auth\_events} — to three separate Snowflake tables.
A naive implementation would launch three separate Flink jobs, each with its own checkpoint cycle,
TaskManager slots, and JDBC connection pool.

Flink's \textbf{Table API \texttt{StatementSet}} pattern allows all three \texttt{INSERT INTO}
statements to be registered and submitted as a \textbf{single job}:

\begin{lstlisting}[language=Python, caption={StatementSet submits three INSERT pipelines as one Flink job.}]
stmt_set = table_env.create_statement_set()
stmt_set.add_insert("sink_listen_events",    listen_query)
stmt_set.add_insert("sink_page_view_events", page_view_query)
stmt_set.add_insert("sink_auth_events",      auth_query)
stmt_set.execute().get_job_client()
\end{lstlisting}

The result is one shared execution graph with one checkpoint barrier cycle covering all three
pipelines simultaneously.
A single checkpoint means the offsets for all three Kafka topics advance atomically — there is
no window of inconsistency where one topic is checkpointed ahead of another.
This is structurally impossible to achieve with three independently running Spark queries.

\subsection{Python Support via PyFlink}

The Streamify implementation is written entirely in Python — a natural choice given the
data engineering team's existing Python proficiency.
PyFlink exposes both the Table API (high-level SQL-like operations) and the DataStream API
(low-level operator control) from Python, and supports Python UDFs for custom transformations.

In Streamify, a Python UDF (\texttt{string\_decode}) is registered directly in the Flink
execution environment to convert latin1-escaped artist and song names to UTF-8 before writing
to Snowflake:

\begin{lstlisting}[language=Python, caption={Registering a Python UDF in PyFlink.}]
@udf(result_type=DataTypes.STRING())
def string_decode(s: str) -> str:
    try:
        return s.encode('latin1').decode('unicode-escape').encode('latin1').decode('utf-8')
    except Exception:
        return s

table_env.create_temporary_function("string_decode", string_decode)
\end{lstlisting}

Spark also supports Python (PySpark), but PySpark UDFs suffer from a significant performance
penalty: each UDF call requires serialising data from the JVM to the Python interpreter via
the Arrow/Pickle bridge.
PyFlink UDFs for the Table API run inside the Python worker process and access state without
round-tripping through the JVM for scalar functions, making them more efficient for
record-level transformations like string decoding.

\subsection{Native Exactly-once Semantics with the JDBC Sink}

Streamify's correctness requirement is strict: \textbf{no event may appear twice in Snowflake},
even if the Flink job is restarted mid-run.
Achieving exactly-once end-to-end with Spark requires the application to implement idempotent
writes (e.g.\ using a unique key to detect and overwrite duplicates) or to use a transactional
sink connector — additional engineering effort beyond the core pipeline logic.

Flink's checkpointing protocol natively integrates with JDBC sinks via a two-phase commit (2PC)
pattern:
\begin{enumerate}
  \item During a checkpoint, the JDBC sink \textbf{pre-commits} the current batch of rows to a
        Snowflake staging area but does not make them visible.
  \item Once the JobManager declares the checkpoint complete, the sink \textbf{finalises} the
        commit, making the rows visible to queries.
  \item If the job crashes before the checkpoint completes, the pre-committed rows are rolled back.
        On restart, Flink replays from the last completed checkpoint offset, and the sink
        re-inserts the same rows — with exactly the same outcome.
\end{enumerate}

This protocol required no application-level deduplication logic in Streamify.
The exactly-once guarantee is enforced at the connector layer, and verified experimentally by
deliberately killing the Flink TaskManager mid-run and confirming that no duplicate rows
appeared in Snowflake after recovery.
