\section{Problem Statement}

A music streaming platform generates three distinct categories of events in real time:

\begin{itemize}
  \item \textbf{Song-play events} (\texttt{listen\_events}) — fired each time a user starts playing a track, carrying metadata such as artist name, song title, duration, user location, and session information.
  \item \textbf{Page-view events} (\texttt{page\_view\_events}) — fired on every navigation action (e.g.\ visiting the home page, search results, or an artist profile), capturing HTTP method, status code, and user context.
  \item \textbf{Authentication events} (\texttt{auth\_events}) — fired on login and logout, recording session validity and subscription level.
\end{itemize}

To illustrate the scale of these events, Table~\ref{tab:event-rates} shows estimated
event volumes under a modest 10,000 concurrent user simulation:

\begin{table}[H]
  \centering
  \caption{Estimated event rates for a 10,000 concurrent user simulation.}
  \label{tab:event-rates}
  \begin{tabularx}{\textwidth}{lXX}
    \toprule
    \textbf{Event type}        & \textbf{Avg.\ rate (events/s)} & \textbf{Volume per hour} \\
    \midrule
    \texttt{listen\_events}    & $\sim$300                       & $\sim$1.1 million \\
    \texttt{page\_view\_events}& $\sim$150                       & $\sim$540,000 \\
    \texttt{auth\_events}      & $\sim$50                        & $\sim$180,000 \\
    \midrule
    \textbf{Total}             & $\sim$500                       & $\sim$1.8 million \\
    \bottomrule
  \end{tabularx}
\end{table}

Processing these events in batch is insufficient for four reasons:

\begin{enumerate}
  \item \textbf{Velocity} — at hundreds of events per second, buffering them for even one hour
        means acting on data that is already stale.
        A song that trended at 9\,pm cannot drive the 9\,pm featured-artists chart if the pipeline
        only runs at midnight.
  \item \textbf{Volume} — 1.8 million events per hour accumulate to over 43 million per day.
        A batch job processing an entire day's events requires significant infrastructure and
        introduces a latency proportional to the window size.
  \item \textbf{Variety} — the three event types have different schemas, require different
        transformations (timestamp normalisation, string decoding, partition column derivation),
        and must eventually be joined with dimension data (users, songs, artists) for analytics.
        Handling schema heterogeneity in batch SQL is cumbersome; a streaming job can apply
        per-topic transformations in parallel.
  \item \textbf{Reliability} — any processing outage must not result in data loss.
        A batch job that crashes mid-run may leave partial results or require a full re-run.
        A streaming pipeline with offset-based checkpointing can resume exactly where it stopped,
        processing only the events that were not yet committed.
\end{enumerate}

The core challenge addressed in this report is therefore:
\textit{how to design and implement a fault-tolerant, low-latency streaming pipeline that
ingests heterogeneous music events, transforms and enriches them in real time, and delivers
clean analytical data to a cloud data warehouse — all without data loss or duplicate records.}
