\section{Report Structure}

The remainder of this report is organised as follows:

\begin{description}
  \item[Chapter 2 — Message Queue: Apache Kafka] \hfill \\
    Introduces Apache Kafka as the event transport backbone of the pipeline.
    It explains why a durable message queue is necessary in distributed systems,
    walks through Kafka's core components (producers, brokers, topics, partitions,
    consumer groups), surveys common industry use cases, and traces Kafka's
    metadata architecture evolution from ZooKeeper to the modern KRaft protocol.

  \item[Chapter 3 — Streaming Processing Engines: Spark vs.\ Flink] \hfill \\
    Provides an in-depth comparison of the two leading open-source stream-processing
    frameworks. The chapter covers Apache Spark's micro-batch and continuous
    processing modes, then Apache Flink's true-streaming dataflow model.
    For each engine, the internal mechanisms — checkpointing, state storage,
    fault tolerance, and delivery semantics — are examined in detail.
    A side-by-side comparison table and a justification for choosing Flink in
    the Streamify project conclude the chapter.

  \item[Chapter 4 — Application: Streamify Pipeline] \hfill \\
    Presents the practical implementation of the Streamify project.
    It describes the full technology stack, the event simulation layer (Eventsim),
    the PyFlink streaming job that reads from Kafka and writes to Snowflake,
    the dbt data models that transform raw events into a star schema, and
    the Power BI dashboard that visualises the resulting business insights.
\end{description}
